---
title: "Forecasting Time Series of Apple Stock"
author: "Akram Nour"
date: "2023-03-06"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Forecasting Time Series of Apple Stock Using Univariate Models

```{r}
library(dplyr)
AAPL <- read.csv("AAPL (1).csv")
AAPL$Date <- as.Date(AAPL$Date)
```

```{r}
library(forecast)

AAPL.ts <- ts(AAPL$Open, frequency = 12,
    start = c(1990))
plot(AAPL.ts, main = "Apple Stock - Monthly Opening Price")
```

### Creating a time series for Monthly Opening price of Apple stock from 1990 to 2023.

# Data

```{r}
AAPL.train <- AAPL.ts[1:384] %>% ts(frequency = 12, start = c(1990,1))
plot(AAPL.train, main = "AAPL Opening Price")

tsdisplay(AAPL.train)

plot(decompose(AAPL.train, type = "multiplicative"))

AAPL$month <- factor(strftime(AAPL$Date,"%b"),levels=month.abb)
boxplot(AAPL$Open~month,AAPL, main = "AAPL Seasonal Boxplot")

rAAPL.ts <- decompose(AAPL.train, type = "multiplicative")$random
plot(rAAPL.ts, main = "Random component of AAPL Obtained by Multiplicative Decomposition")

tsdisplay(rAAPL.ts)
```

### I'm going to set the training set to only include values from 1990 to 2022 so that we can predict monthly opening price from 2022 - 2023 (test). Looking at out the multiplicative decomposition, we can notice a clear upward trend and insignificant seasonality. Looking at the seasonal boxplot, because all months have a similar median and variance, we can confirm that seasonality is not present. Looking at the random component, its ACF starts off with serial autocorrelation in lags 1 through 7 but quickly tapers off. Overall, it looks stationary! Now we can use our univariate models to forecast opening prices.

# ARIMA Model

```{r}
library(ggplot2)

model <- arima(AAPL.train, order = c(0,1,0))
tsdisplay(model$residuals[-1])

AIC(model)

pred <- predict(model, n.ahead = 13)
model.predicted <- pred$pred
model.predicted.se <- pred$se
AAPL.Test.real <- AAPL.ts[384:396]
AAPL.Test.real


ggplot(AAPL, aes(x=1:396)) + 
  geom_line(aes(y=Open)) +
  geom_line(data=data.frame(model.predicted), aes(x=c(384:396), y = model.predicted, color = "red")) +
  xlab("Time") + ggtitle("ARIMA Forecast") + theme(legend.position = "none")
```

# ARIMA + GARCH Model

```{r}
acf((model$residuals[-1])^2,
    main = "ARIMA(0,1,0) Residuals^2")

library(fGarch)
library(tseries)

garch.3 <- garch(model$residuals, order = c(1,1), trace = F)
AIC(garch.3)

garch.4 <- garch(model$residuals, order = c(0,2), trace = F)
AIC(garch.4)

help("garchFit")

d1 <- diff(AAPL.train, differences = 1)
garch.fit <- garchFit(formula = ~arma(0,0)+garch(1,1), data = d1,
                      trace = F)
garch.pred <- predict(garch.fit, n.ahead = 12, trace = FALSE, mse = c("cond"))
garch.forecast <- ts(garch.pred[,1]+garch.pred[,2])

garch.fc.og <- c()
for(i in 1:13){
  garch.fc.og[i] <- AAPL.train[384] + 
    sum(garch.forecast[1:i])
}

ggplot(AAPL, aes(x = 1:396)) + geom_line(aes(y=Open)) + geom_line(data = data.frame(garch.fc.og), aes(x=c(384:396),
                                                                                                      y=garch.fc.og, color = "red")) + 
  xlab("Time") + ggtitle("ARIMA+GARCH Forecast") + theme(legend.position = "none")
```

# Exponential Smoothing Model

```{r}
hw1 <- HoltWinters(AAPL.train, gamma = FALSE)

hw1$alpha
hw1$beta

hw1.pred <- predict(hw1, n.ahead = 13)

hw.table1 <- cbind(as.data.frame(AAPL.Test.real),as.data.frame(hw1.pred)) %>% as.data.frame() %>% dplyr::select("observed" = 1, "predicted" = 2)

ggplot(AAPL, aes(x=1:396)) + geom_line(aes(y=Open)) + geom_line(data=data.frame(hw1.pred),  aes(x=c(384:396), y=hw1.pred, color = "red")) + xlab("Time") + ggtitle("Holt-Winters Forecast") + theme(legend.position = "none")

```
